Realizing Temporal Graphs

### 2023-01-29 RealizingTemporalGraphs.tex
Changes Saturday 28th January

@@

Observe that edges incident with $y^T, y^F$ that have label three connect $y^T, y^F$ to $v$. Furthermore, $v$ has distance two from $x$ in the undetermined graph $G$. Since all edges incident with variable vertices $x$ have label two, there is no temporal path starting at $x$ and arriving at $v$ at time two, that could then continue to $y_T$ or $y_F$ using an edge with label three. Hence, there is no temporal path from $x$ to $y^T, y^F$ using an edge with label three. Since all edges incident with $x$ have label two and variable vertices $x$ are non-adjacent to variable vertices $y^T,y^F$, there cannot be a temporal path from variable vertex $x$ to variable vertex $y^T$ or $y^F$.
\item For all variable vertices $x$ and clause vertices $c$ we have $d(v,x)=d(v,c)=\infty$:
\todo[inline]{New matrix entries D for period Delta: d(v,x)=Delta (using two edges, with labels L+1,L). 
New matrix entries D for period Delta: d(v,c)=Delta (using two edges, with labels L,L-1).}

@@

    For each color $i$, we introduce the following vertices. What we describe in the following will be used as a \emph{verification gadget} for color $i$.
    \begin{itemize}
        \item We have $k-1$ vertices $v^i_\ell$ for $\ell\in[k]\setminus\{i\}$ and one vertex $v^i_0$.
        \item If $1\neq i\neq k$, then we have a set $\hat{U}_i$ of $n$ additional vertices $\hat{u}^i_1,\hat{u}^i_2,\ldots,\hat{u}^i_n$.
    \end{itemize}
    We add the following edges. 
    %Assume $1\neq i \neq k$. We describe the two cases $i=1$ and $i=k$ afterwards. 
    If $i\neq 1$, let $e_\ell\in F_{1,i}$ let $w^i_{\ell'}\in W_i$ be incident with $e_\ell$. Then we add edge $\{v^i_0,u^\ell_{\ell'}\}$ between $v^i_0$ and the vertex $u^\ell_{\ell'}$ of the edge selection gadget of color combination $1,i$.

    Let $1<j<i$ (skip if $i=1$) and let $e_\ell\in F_{j,i}$ let $w^i_{\ell'}\in W_i$ be incident with $e_\ell$. Then we add edge $\{v_{j-1}^i,u^\ell_{\ell'}\}$ between $v^i_{j-1}$ and the vertex $u^\ell_{\ell'}$ of the edge selection gadget of color combination $j,i$.
    Furthermore, we add edge $\{v_{j}^i,u^\ell_{\ell'+1}\}$ between $v^i_j$ and the vertex $u^\ell_{\ell'+1}$ of the edge selection gadget of color combination $j,i$.

    We add edge $\{v^i_{i-1},\hat{u}^i_1\}$ and for all $\ell\in[n-1]$ we add edge $\{\hat{u}^i_\ell,\hat{u}^i_{\ell+1}\}$.

    Let $e_\ell\in F_{i,i+1}$ let $w^i_{\ell'}\in W_i$ be incident with $e_\ell$. Then we add edge $\{\hat{u}^i_n,u^\ell_{n+\ell'}\}$ between $\hat{u}^i_n$ and the vertex $u^\ell_{n+\ell'}$ of the edge selection gadget of color combination $i,i+1$.

    Let $i<j\le k$ (skip if $i=k$) and let $e_\ell\in F_{i,j}$ let $w^i_{\ell'}\in W_i$ be incident with $e_\ell$. Then we add edge $\{v_{j-1}^i,u^\ell_{n+\ell'}\}$ between $v^i_{j-1}$ and the vertex $u^\ell_{n+\ell'}$ of the edge selection gadget of color combination $i,j$.
    Furthermore, we add edge $\{v_{j}^i,u^\ell_{n+\ell'+1}\}$ between $v^i_j$ and the vertex $u^\ell_{n+\ell'+1}$ of the edge selection gadget of color combination $i,j$.

@@

This finished the description of the underlying graph $G$. We can observe that the vertex set containing
\begin{itemize}
    \item vertices $v_{i,j}^\star$ and $v_{i,j}^{\star\star}$ of each edge selection gadget,
    \item vertices $v^i_\ell$ with $\ell\in\{0,1,\ldots,i-1,i+1,\ldots,k\}$ and vertex $\hat{u}^i_n$ (if exists) of each validation gadget.
    \item vertices $\hat{v}_1$ and $\hat{v}_2$ of each connector gadget, and
    \item vertex $w^\star$ of the alignment gadget
\end{itemize}
forms a feedback vertex set in $G$ with size $O(k^2)$.

@@

Next, consider the edge selection gadget for color combination $i,j$ with $i<j$.
\end{proof}

\section{FPT and W[1]-hardness of exact realization -- new version}
In a graph $G=(V,E)$, a \emph{feedback edge set} $F \subset E(G)$ is a subset of edges, such that each cycle in $G$ has at least one edge in $F$.
The minimum such set $F$ is called a  \emph{minimum feedback edge set} and its size, $k = |F|$, is called the \emph{feedback edge number} of graph $G$.
Note that one can find a minimum feedback edge set in linear time, by calculating a spanning tree (or forest) $T$ of the given graph $G$ and then removing all of the edges $T$ from $G$, \ie $F = E(G) \setminus E(T)$.

@@

For an easier understanding and the analysis of the algorithm we give the following definition.
\begin{definition}
    Let $U \subseteq V(G')$ be a set of vertices of interest and let $u,v \in U$.
    A path $P = (u=v_1,v_2, \dots, v_p = v)$  in graph $G'$, where all inner vertices are not in $U$, \ie $v_i \notin U$ for $i \in \{ 2, 3, \dots, p-1\}$,
    is called a \emph{segment} from $u$ to $v$. We denote it as $S_{u,v}$.
    Note $S_{u,v} \neq S_{v,u}$.
\end{definition}
From the definition it follows, that there are $O(k^2)$ segments in $G'$.
Observe also that a temporal path in $G'$ between two vertices of interest is either a segment, or consists of a sequence of some segments.
%For a pair $u,v$ of vertices in $U$, there are $O(k!)=O(k^k)$ possible paths in $G'$ between them.

@@

    \item The fastest temporal paths among all pairs of vertices of $U$.

@@

    \item The fastest temporal paths among all vertices in $U^*$, 

@@

    \item The fastest temporal paths from vertices of $U$ to vertices in $U^*$,

@@

    \item Let $S_{u,v} = (u=v_1,v_2, \dots, v_p = v)$ and $S_{w,z} = (w=z_1,z_2, \dots, z_r = z)$ be two segments,

@@

    there are $O(k^{k^5})$ possible paths of this form. 

@@

    \item \label{FPT:guess-uToSegment}

@@

    $w \leadsto u \rightarrow v_2$, $w \leadsto v \rightarrow v_2$,

@@

    $v_2 \rightarrow u \leadsto w$, $v_2 \rightarrow v \leadsto w$.

@@

@@

    $D_{v_i,z_{p-1}} < D_{v_i,z_{p-2}} < \dots < D_{v_i,z_j}$.

@@

@@

    Besides that we want to guess also hot the fastest temporal paths from $v_i$ to $z_{j-1}, z_{j+1}$ start and end.

@@

    Then it holds that $D_{v_i,z_2} < D_{v_i,z_3} < \dots < D_{v_i,z_j} = D_{v_i,z_{j+1}} > D_{v_i,z_{j + 2}} > \cdots > D_{v_i,z_{p-1}}$ and 
    $D_{v_i,z_{p-1}} < D_{v_i,z_{p-2}} < \dots < D_{v_i,z_j}$.

@@

    In this case we only need to guess the following two fastest temporal paths $u \leadsto w \rightarrow z_2$

@@

    Note that in this case we do not introduce any new guesses, as we guessed these paths already in Step~\ref{FPT:guess-uToSegment}.

@@

We create all of the guesses independently.
We start with the first one, that results in $O(k^{k^3})$ different possibilities, then we split each one of these guesses into $O(k^{k^6})$ new ones, that respond to the guessing in the second step, etc.
After creating all of the guesses we end up with exponential in $k$ different cases.
We now create one partial ILP instance for each case.
From now on we assume that we know exactly the structure of all paths we have guessed and the permutation of all variables.

@@

\begin{proof}
We denote with $v_i \in U$ a vertex of interest in $P \setminus \{u,v\}$.
There are two options, either $v_i$ is reached from $u$ using the subpath $P_i = (u, v_2, v_3, \dots, v_i)$ of $P$,
or there exists a fastest temporal path $P' = (u, w_2, w_3, \dots, w_{p'} = v_i)$, for which $P \cap P' = \{u, v_i\}$.
If the temporal path from $u$ to $v_i$ is a subpath of $P$, then we determine the labeling of $P$ using \cref{claim:FPT-uv-Labelalledges}.
So we can suppose that the fastest temporal path from $u$ to $v_i$ is of form $P'$.\todo{TBD - George proved it}
\end{proof}

@@

\subsection{Adding constraints and variables to the ILP}
First case we observe is the case when we want to determine the labels on fastest temporal paths among vertices of interest.
We proceed in the following way.
Let $u,v \in U$ be two vertices of interest and let $P_{u,v}$ be the fastest temporal path from $u$ to $v$.
If $P_{u,v}$ is a segment we determine all the labels of edges of $P$, with respect to the label of the first edge (see \cref{claim:FPT-uv-Labelalledges}).
%Note that since we determined labels of all edges, with respect to the first edge, we also know the exact values of waiting times at each inner vertex of $P_{u,v}$, when traversing the path.
In the case when $P{_u,v}$ is a sequence of $\ell$ segments, we determine all but $\ell - 1$ labels of edges of $P$, with respect to the label of the first edge (see \cref{claim:FPT-uv-LabelAlmostalledges}).
%In both cases we introduce the equality constraint for $P_{u,v}$ (as in \cref{eq:FPT-equalityConstraint}).
%Since there are $\ell - 1$ edge labels that are not determined with respect to the label of the first edge, may be some  TODO from above??

After finishing the previous step for all pairs of vertices in $U$,
the edges of fastest temporal paths from $u$ to $v$, where $u,v \in U$, are determined with respect to the label of the first edge of each path.
Let us call these edges \emph{determined edges}, and the edges that were not determined with the respect to some other edge label up to now the \emph{undetermined} edges.
If the fastest temporal path among two vertices $u,v \in U$ is just an edge $e$, then we treat it as being determined, since it gets assigned a label $\lambda(e)$ with respect to itself.
Note that the undetermined edges are exactly the ones that are not a part of any fastest temporal path.
Let us first observe some properties of such edges.

\begin{claim}
    Undetermined edges form disjoint segments in $G'$, where each segment starts and ends in some vertex of interest, 
    \ie for every undetermined path $P$ there exist $u,v \in U$, such that $P = (u, v_1, v_2, \dots, v_p, v)$, where $p \geq 0$ and $v_1,v_2,\dots,v_p \in V(G') \setminus U$.
\end{claim}
\begin{proof}
    We prove this claim in two steps.
    First we prove that all inner vertices of $P$, \ie vertices $v_i \in  S \setminus \{u,v\}$ are of degree exactly $2$ and are not vertices of interest. 
    They are of degree at least $2$ since they are on a path $P$ and are incident to two undetermined edges in  $P$.
    They are of degree at most $2$ as if they were of a higher degree they would be vertices of interest and in this case we could split $P$ into shorter segments (subpaths).

    Second, suppose that $P$ does not start in a vertex of interest, therefore $u \notin U$. As $u$ is the starting vertex of the undetermined path $P$, there must exist a vertex of interest $u'$ such that a path between $u$ and $u'$ is labeled.
    Since $u$ is not a vertex of interest, the only reason to label edges between $u'$ and $u$ would be if they were a part of some fastest temporal path among two vertices of interest and $u$ would be an inner vertex on that path.
    Therefore $u$ would have to be incident to two labeled edges (that are part of some fastest temporal path) and one undetermined edge, but in this case the degree of $u$ would be at least $3$ and therefore $u$ would be a vertex of interest.

    From all of the above, it follows that $P$ is an undetermined segment in $G'$.
\end{proof}

\begin{claim}
    After determining the labels of fastest temporal paths between all pairs of vertices of interest,
    there can be at most one undetermined segment among each pair of them.
    %among any pair of vertices of interest $u,v \in U$ there can be at most one undetermined segment $S = (u, v_1, v_2, \dots, v_p, v)$, where $p \geq 0$ and $v_1,v_2,\dots,v_p \in V(G') \setminus U$.
\end{claim}
\begin{proof}
   Suppose for the contradiction that after determining the labels of all fastest temporal paths between vertices of interest,
   there are two undetermined segments among $u,v \in U$.
   Denote them with $S_{u,v}$ and $S'_{u,v}$. 
   Then $C = S_{u,v} \cup S_{u,v}'$ forms a cycle in $G'$. Since $G'$ admits no loops, the cycle $C$ is of length at least $3$.
   Since $C$ is a cycle, at least one of its edges $e \in E(C)$ has to be in the feedback edge set $F$, and therefore, by the definition of vertices of interest $U$, 
   both endpoints of $e$ have to be in $U$.
   If $e = uv$ then there are only two vertices of interest in $C$, but the label of edge $e$ is determined.
   And if $e \neq uv$ then there are at least $3$ vertices of interest in $C$.
   So $S_{u,v}$ and $S_{u,v}'$ cannot be both fully undetermined.
\end{proof}

%From the above it follows that there are at most $O(|U|^2) = O(k^2)$ undetermined segments in $G'$ after the algorithm finishes with determining the labels for all fastest temporal paths among vertices of interest.
Now we want to relate the undetermined segments with the determined ones.
Let $S_{u,v}$ and $S_{w,z}$ be two segments.
At the beginning we have guessed what is a split vertex of a fixed vertex $v_i \in S_{u,v}$ in the segment $S_{w,z}$. 
Denote this vertex as $z_j$, denote its neighbor closer to $w$ in $S_{w,z}$ as $z_{j-1}$,
and its neighbor closer to $z$ in $S_{w,z}$ as $z_{j+1}$.
We also guessed how the fastest temporal path from $v_i$ to $z_j$ starts and ends, which determined also the structure of the fastest temporal paths 
from $v_i$ to $z_{j-1}$ and from $v_i$ to $z_{j+1}$
(for the details see guess G-\ref{FPT:guess-splitFromAnotherSegmentAndPaths}). \todo{Two cases.}
W.l.o.g. assume that we have guessed that the fastest path from $v_i$ to $z_j$
passes through $w$ and $z_{j-1}$.
Then the fastest temporal path from $v_i$ to $z_{j+1}$ passes through $z$.
And all fastest temporal paths from $v_i$ to any $z_{j'} \in S_{w,z}$
use all of the edges in $S_{w,z}$ with the exception of the edge $z_j z_{j+1}$.
Using this information and \cref{obs:FirstLabelAndDuration}, we can determine the labels on all edges, with respect to the first or last label from the segment $S_{u,v}$,
with the exception of the edge $z_j z_{j+1}$.
Therefore, all edges of $S_{w,z}$ but $z_j z_{j+1}$ become determined.
Since we repeat that procedure for all pairs of segments,
we get that for a fixed segment $S_{w,z}$ we end up with an undetermined edge $z_j z_{j+1}$
if and only if, this is an undetermined edge in relation to every other segment $S_{u,v}$ and its fixed vertex $v_i$.

>>>>

\newcounter{guesscounter}

@@

Observe that edges incident with $y^T, y^F$ that have label three connect $y^T, y^F$ to $v$. Furthermore, $v$ has distance two from $x$ in the underlying graph $G$. Since all edges incident with variable vertices $x$ have label two, there is no temporal path starting at $x$ and arriving at $v$ at time two, that could then continue to $y_T$ or $y_F$ using an edge with label three. Hence, there is no temporal path from $x$ to $y^T, y^F$ using an edge with label three. Since all edges incident with $x$ have label two and variable vertices $x$ are non-adjacent to variable vertices $y^T,y^F$, there cannot be a temporal path from variable vertex $x$ to variable vertex $y^T$ or $y^F$.
\item For all variable vertices $x$ and clause vertices $c$ we have $d(v,x)=d(v,c)=\infty$:
\todo[inline]{New matrix entries D for period Delta: d(v,x)=Delta (using two edges, with labels L+1,L). 
New matrix entries D for period Delta: d(v,c)=Delta (using two edges, with labels L,L-1).}

@@

    For each color $i$, we introduce the following vertices. What we describe in the following will be used as a \emph{verification gadget} for color $i$.
    \begin{itemize}
        \item We have $k-1$ vertices $v^i_\ell$ for $\ell\in[k]\setminus\{i\}$ and one vertex $v^i_0$.
        \item If $1\neq i\neq k$, then we have a set $\hat{U}_i$ of $n+3$ additional vertices $\hat{u}^i_1,\hat{u}^i_2,\ldots,\hat{u}^i_{n+3}$.
    \end{itemize}
    We add the following edges. 
    %Assume $1\neq i \neq k$. We describe the two cases $i=1$ and $i=k$ afterwards. 
    %If $i\neq 1$, let $e_\ell\in F_{1,i}$ let $w^i_{\ell'}\in W_i$ be incident with $e_\ell$. Then we add edge $\{v^i_0,u^\ell_{\ell'}\}$ between $v^i_0$ and the vertex $u^\ell_{\ell'}$ of the edge selection gadget of color combination $1,i$.

    Let $1\le j<i$ (skip if $i=1$), let $e_\ell\in F_{j,i}$, and let $w^i_{\ell'}\in W_i$ be incident with $e_\ell$. Then we add edge $\{v_{j-1}^i,u^\ell_{\ell'}\}$ between $v^i_{j-1}$ and the vertex $u^\ell_{\ell'}$ of the edge selection gadget of color combination $j,i$.
    Furthermore, we add edge $\{v_{j}^i,u^\ell_{\ell'+1}\}$ between $v^i_j$ and the vertex $u^\ell_{\ell'+1}$ of the edge selection gadget of color combination $j,i$.

    If $1\neq i\neq k$, then we add edge $\{v^i_{i-1},\hat{u}^i_1\}$ and for all $\ell\in[n+2]$ we add edge $\{\hat{u}^i_\ell,\hat{u}^i_{\ell+1}\}$. 
    Let $e_\ell\in F_{i,i+1}$ and let $w^i_{\ell'}\in W_i$ be incident with $e_\ell$. If $i\neq 1\neq k$, then we add edge $\{\hat{u}^i_{n+3},u^\ell_{n+\ell'}\}$ between $\hat{u}^i_{n+3}$ and the vertex $u^\ell_{n+\ell'}$ of the edge selection gadget of color combination $i,i+1$. If $i=1$, then we add edge $\{v^i_0,u^\ell_{n+\ell'}\}$ between $v^i_0$ and the vertex $u^\ell_{n+\ell'}$ of the edge selection gadget of color combination $i,i+1$.

    Let $i<j\le k$ (skip if $i=k$), let $e_\ell\in F_{i,j}$, and let $w^i_{\ell'}\in W_i$ be incident with $e_\ell$. Then we add edge $\{v_{j-1}^i,u^\ell_{n+\ell'}\}$ between $v^i_{j-1}$ and the vertex $u^\ell_{n+\ell'}$ of the edge selection gadget of color combination $i,j$.
    Furthermore, we add edge $\{v_{j}^i,u^\ell_{n+\ell'+1}\}$ between $v^i_j$ and the vertex $u^\ell_{n+\ell'+1}$ of the edge selection gadget of color combination $i,j$.

@@

This finished the description of the underlying graph $G$. We can observe that the vertex set containing
\begin{itemize}
    \item vertices $v_{i,j}^\star$ and $v_{i,j}^{\star\star}$ of each edge selection gadget,
    \item vertices $v^i_\ell$ with $\ell\in\{0,1,\ldots,i-1,i+1,\ldots,k\}$ and vertex $\hat{u}^i_{n+3}$ (if exists) of each validation gadget.
    \item vertices $\hat{v}_1$ and $\hat{v}_2$ of each connector gadget, and
    \item vertex $w^\star$ of the alignment gadget
\end{itemize}
forms a feedback vertex set in $G$ with size $O(k^2)$.

@@

Next, consider the edge selection gadget for color combination $i,j$ with $i<j$.
\begin{itemize}
    \item Let $1\le \ell<\ell'\le m$. We set $d(x_\ell,x_\ell')=2(i+j)\cdot(\ell'-\ell)$.
    \item For all $\ell\in[m]$ we set $d(x_\ell,v_{i,j}^{\star\star})=4n+4$.
\end{itemize}

Next, consider the validation gadget for color $i$. First, assume $1\neq i \neq k$. Let $1\le j<i$, let $e_\ell\in F_{j,i}$, and let $w^i_{\ell'}\in W_i$ be incident with $e_\ell$. We set the following.
\begin{itemize}
    \item We set $d(v^i_{j-1},w^i_{\ell'+1})=3$.
    %\item We set $d(w^i_{\ell'},v^i_{j})=3$.
\end{itemize}

For all $0\le j<j'<i$ we set the following.
\begin{itemize}
    \item We set $d(v^i_j,v^i_{j'})=6(j'-j)$.
\end{itemize}
\end{proof}

\section{FPT}
In a graph $G=(V,E)$, a \emph{feedback edge set} $F \subset E(G)$ is a subset of edges, such that each cycle in $G$ has at least one edge in $F$.
The minimum such set $F$ is called a  \emph{minimum feedback edge set} and its size, $k = |F|$, is called the \emph{feedback edge number} of graph $G$.
Note that one can find a minimum feedback edge set in linear time, by calculating a spanning tree (or forest) $T$ of the given graph $G$ and then removing all of the edges $T$ from $G$, \ie $F = E(G) \setminus E(T)$.

@@

For an easier understanding and the analysis of the algorithm we give the following definition.
\begin{definition}
    Let $U \subseteq V(G')$ be a set of vertices of interest and let $u,v \in U$.
    A path $P = (u=v_1,v_2, \dots, v_p = v)$  in graph $G'$, where all inner vertices are not in $U$, \ie $v_i \notin U$ for $i \in \{ 2, 3, \dots, p-1\}$,
    is called a \emph{segment} from $u$ to $v$. We denote it as $S_{u,v}$.
    Note $S_{u,v} \neq S_{v,u}$.
\end{definition}
From the definition it follows, that there are $O(k^2)$ segments in $G'$. \todo{True but does not follow directly just from the definition}
Observe also that a temporal path in $G'$ between two vertices of interest is either a segment, or consists of a sequence of some segments.
%For a pair $u,v$ of vertices in $U$, there are $O(k!)=O(k^k)$ possible paths in $G'$ between them.

@@

    \item \label{FPT-guessFTPamongU}
    The fastest temporal paths among all pairs of vertices of $U$.

@@

    \item \label{FPT-guessFTPamongUstar}
    The fastest temporal paths among all vertices in $U^*$, 

@@

    \item \label{FPT-guessFTPamongUandUstar}
    The fastest temporal paths from vertices of $U$ to vertices in $U^*$,

@@

    \item \label{FPT-guessFTPamongv2z2}
    Let $S_{u,v} = (u=v_1,v_2, \dots, v_p = v)$ and $S_{w,z} = (w=z_1,z_2, \dots, z_r = z)$ be two segments,

@@

    there are $O(k^{k^5})$ possible paths of this form. \\

@@

    \item \label{FPT:guess-uToSegmentz2}

@@

    $w \leadsto u \rightarrow v_2$, $w \leadsto v \rightarrow v_{p-1} \rightarrow \cdots \rightarrow v_2$,

@@

    $v_2 \rightarrow u \leadsto w$, $v_2 \rightarrow v_3 \rightarrow \cdots v \leadsto w$.

@@

    In the case when $S_{u,v}$ is of length $4$, the fixed vertex $v_i$ is the middle vertex (\ie vertex that is on distance two from $u$ and $v$ in $S_{u,v}$).
    If $S_{u,v}$ is not of length $4$, we fix an arbitrary vertex $v_i \in S_{u,v}$.

@@

    $D_{v_i,z_{r-1}} < D_{v_i,z_{r-2}} < \dots < D_{v_i,z_j}$.

@@

    the fastest temporal paths from $v_i$ to $z_{r-1}, z_{r-2}, \dots, z_{j+1}$ go through $z$.

@@

    Besides that we want to guess also how the fastest temporal paths from $v_i$ to $z_{j-1}, z_{j+1}$ start and end.

@@

    Then it holds that $D_{v_i,z_2} < D_{v_i,z_3} < \dots < D_{v_i,z_j} = D_{v_i,z_{j+1}} > D_{v_i,z_{j + 2}} > \cdots > D_{v_i,z_{p-1}}$.

@@

    In this case we only need to guess the following two fastest temporal paths 
    $u \leadsto w \rightarrow z_2$

@@

    Note that in this case we do not introduce any new guesses, as we have aready guessed 
    the fastest paths of form
    $u \leadsto w \rightarrow z_2$
    and $u \leadsto z \rightarrow z_{p-1}$
    (see guess~G-\ref{FPT:guess-uToSegmentz2}).

@@

    \item \label{FPT:guess-splitFromUtoAnotherSegment}
    Let $w \in U$ be a vertex of interest and let $S_{u,v} = (u=v_1,v_2, \dots, v_p = v)$ be a line segment in $G'$.
    Similarly as above we want to guess a split vertex of $w$ in $S_{u,v}$, and the fastest temporal path that reaches it.
    We again have two cases,
    first one where $v_i$ is a unique vertex in $S_{u,v}$ that is furthest away from $w$,
    and the
    second one where $v_i, v_{i+1}$ are two incident vertices in $S_{u,v}$, that are furthest away from $w$.
    In first case we know exactly how the fastest paths from $w$ to all vertices $v_j \in S_{u,v} \setminus \{v_i\}$
    travel through the segment $S_{u,v}$ (\ie either through $u$ or $v$).
    Therefore we have to guess how the fastest path from $w$ reaches vertex $v_i$,
    we have two options, either it travels through $u \rightarrow v_2 \rightarrow \cdots \rightarrow v_{i-1} \rightarrow v_i$
    or 
    $v \rightarrow v_{p-1} \rightarrow \cdots \rightarrow v_{i+1} \rightarrow v_i$.
    Which produces two new guesses.
    In the second case we know exactly how the fastest temporal path reaches $v_i$ and $v_{i+1}$, and consequently all the inner vertices.
    Therefore no new guesses are needed.
    Note that the above guesses, together with the guesses from G-\ref{FPT:guess-uToSegmentz2},
    uniquely determine fastest temporal paths from $w$ to all vertices in $S_{u,v}$ 
    (this also holds for the case when $w \in S_{u,v}$, \ie $w = u$ or $w = v$).

    All together we make two guesses for each pair of vertex $w \in U$ and segment $S_{u,v}$.
    We repeat this for all vertices of interest, and all segments,
    which produces $O(k^{k^2})$ new guesses.
    % counter for guesses continues
    \setcounter{guesscounter}{\value{enumi}}
\end{enumerate}
We create all of the guesses independently.
We start with the first one, that results in $O(k^{k^3})$ different possibilities, then we split each one of these guesses into $O(k^{k^6})$ new ones, that respond to the guessing in the second step, etc.
After creating all of the guesses we end up with exponential in $k$ different cases.
We now create one partial ILP instance for each case.
From now on we assume that we know exactly the structure of all paths we have guessed and the permutation of all variables.

@@

\begin{proof}
We denote with $v_i \in U$ a vertex of interest in $P \setminus \{u,v\}$.
There are two options, either $v_i$ is reached from $u$ using the subpath $P_i = (u, v_2, v_3, \dots, v_i)$ of $P$,
or there exists a fastest temporal path $P' = (u, w_2, w_3, \dots, w_{p'} = v_i)$, for which $P \cap P' = \{u, v_i\}$.
If the temporal path from $u$ to $v_i$ is a subpath of $P$, then we determine the labeling of $P$ using \cref{claim:FPT-uv-Labelalledges}.
So we can suppose that the fastest temporal path from $u$ to $v_i$ is of form $P'$.
\todo[inline]{George proved it. Also: add notation that is used in the proof to the beginning of the section, namely: $d(u,v,P)$ is the duration of the temporal path from $u$ to $v$ that traverses the edges of $P$.}
\end{proof}

@@

\begin{lemma}\label{lemma:FPT-noUndeterminedEdgeInSegment}
Let $S_{u,v}$ be a segment in $G$ of length at least $5$,
\ie $S_{u,v}= (u=v_1,v_2, \dots, v_p = v)$, where $p > 5$.
It cannot happen that an inner edge $f = v_i v_{i+1}$ from $S_{u,v} \setminus \{u,v\}$,
is not a part of any fastest temporal path, of length at least $2$, among vertices in $S_{u,v}$,
\ie there has to be a pair $v_j, v_{j'} \in S_{u,v}$ \st the fastest temporal path from $v_j$ to $v_{j'}$ passes through edge $f$.
In the case when $p = 5$ all temporal paths of length $2$ avoid $f$ if and only if $f$ has the same label as both of the edges incident to it.
\end{lemma}

\begin{proof}
    For an easier understanding and better readability we present the proof for $S_{u,v}$ of fixed length $5$.
    The case, when $S_{u,v}$ is longer easily follows from presented results.

    Let $S_{u,v} = (u=v_1,v_2, v_3, v_4, v_5, v_6=v)$.
    We distinguish two cases, first when $f = v_2v_3$ (note that the case with $f = v_4v_5$ is symmetrical),
    and the second when  $f = v_3v_4$.
    Throughout the proof we denote with $t_i$ the label of edge $v_i v_{i+1}$.
    Suppose for the contradiction, that none of the fastest temporal paths among vertices of $S_{u,v}$ traverses the edge $f$.

    Case 1,  $f = v_2v_3$.
    Let us observe the case of fastest temporal paths among $v_{1}$ and $v_{3}$.
    Denote with $Q = (v_{1}, v_2, v_3)$ and with $P' = (v_3,v_4,v_5,v_6)$.
    From our proposition it follows that
    \begin{itemize}
        \item the fastest temporal path $P ^ +$ from $v_1$ to $v_3$ 
    is of the following form 
    $P^+ = v_{1}  \leadsto v_6 \rightarrow v_5 \rightarrow v_4 \rightarrow v_3$,
    and
        \item the fastest temporal path $P ^ -$ from $v_{3}$ to $v_{1}$ 
    is of the following form 
    $P^- = v_{3} \rightarrow v_{4} \rightarrow v_5 \rightarrow v_6 \leadsto v_1$.
    \end{itemize}
    It follows that 
    $
    d(v_{1}, v_{3}, P^+) \leq d(v_{1}, v_{3}, Q),
    $
    and
    $
    d(v_{1}, v_{3}, P^-) \leq d(v_{1}, v_{3}, Q)
    $.
    Let 
    Note that $d(v_{1}, v_{3}, P^+) \geq 1 + d(v_6,v_3,P')$,
    and by the definition $d(v_6,v_3,P') = 1 + (t_4 - t_5)_\Delta + (t_3 - t_4)_\Delta$,
    where $(t_i - t_j)_\Delta$ denotes the difference of two consecutive labels $t_i, t_j$ modulo $\Delta$.
    Similarly holds for $d(v_{1}, v_{3}, P^+)$.
    Summing now both of the above equations we get
    \begin{equation}
    \begin{split} \label{eq:FPT-prf-fUnlabeled}
        d(v_{1}, v_{3}, P^+) + d(v_{3}, v_{1}, P^-) &\leq 
        d(v_{1}, v_{3}, Q) + d(v_{3}, v_{1}, Q) \\
        %
        1 + d(v_6,v_3,P') + 1 + d(v_3,v_6,P') &\leq d(v_{1}, v_{3}, Q) + d(v_{3}, v_{1}, Q) \\
        %
        2 + 
        1 + (t_4 - t_5)_\Delta + (t_3 - t_4)_\Delta +
        1 + (t_4 - t_3)_\Delta + (t_5 - t_4)_\Delta 
        &\leq 
        1 + (t_2 - t_1)_\Delta + 
        1 + (t_1 - t_2)_\Delta\\
        %
        (t_4 - t_5)_\Delta + (t_5 - t_4)_\Delta +
        (t_4 - t_3)_\Delta + (t_3 - t_4)_\Delta  + 2 
        &\leq 
        (t_2 - t_1)_\Delta + (t_1 - t_2)_\Delta.
    \end{split}
    \end{equation}
    Note that if $t_i \neq t_j$ we get that 
    the sum
    $(t_i - t_j)_\Delta + (t_j - t_i)_\Delta$ equals exactly $\Delta$,
    and if $t_i = t_j$ the sum equals $2\Delta$.
    This follows from the definition of travel delays at vertices (see \cref{obs:travel-delays-both-directions}).
    Therefore we get from \cref{eq:FPT-prf-fUnlabeled}, 
    that the right part is at most $2 \Delta$, while the left part is at least $2 \Delta + 1$,
    for any relation of labels $t_1,t_2, \dots, t_5$,
    which is a contradiction.

    Case 2,  $f = v_3v_4$.
    Here we consider the fastest paths among vertices $v_{2}$ and $v_{4}$.
    By similar arguments as above we get
    \begin{align*}
        (t_5 - t_1)_\Delta + (t_4 - t_5)_\Delta + (t_5 - t_4)_\Delta + (t_1 - t_5)_\Delta + 2 \leq (t_3 - t_2)_\Delta + (t_2 - t_3)_\Delta,
    \end{align*}
    which is impossible.

    In the case when $S_{u,v}$ is longer, we would get even bigger number on the left hand side of \cref{eq:FPT-prf-fUnlabeled}, 
    so we conclude that in all of the above cases, it cannot happen that all fastest paths of length $2$, among vertices in $S{_u,v}$, avoid an edge $f$.

    Let us observe now the case when $S_{u,v} = (u=v_1,v_2, v_3, v_4, v_5=v)$ is of length $4$.
    Let $f = v_2 v_3$ (the case with $f = v_3 v_4$ is symmetrical).
    Suppose that the fastest temporal paths among $v_1$ and $v_3$ do not use the edge $f$.
    We denote with $R^+$ the fastest path from $v_1$ to $v_3$, 
    which is of form $u \leadsto v \rightarrow v_4 \rightarrow v_3$,
    and similarly 
    with $R^-$ the fastest path from $v_3$ to $v_1$, which is
    of form $v_3 \rightarrow v_4 \rightarrow v \leadsto u$.
    We denote with $R' = (v_3, v_4, v_5)$ and with $S = (v_1,v_2,v_3)$.
    Again we get the following.
    \begin{equation}
    \begin{split}
        d(v_{1}, v_{3}, R^+) + d(v_{3}, v_{1}, R^-) &\leq 
        d(v_{1}, v_{3}, S) + d(v_{3}, v_{1}, S) \\
        %
        1 + d(v_5,v_3,R') + 1 + d(v_3,v_5,R') &\leq d(v_{1}, v_{3}, S) + d(v_{3}, v_{1}, S) \\
        %
        (t_3 - t_4)_\Delta + (t_4 - t_3)_\Delta + 2 
        &\leq 
        (t_2 - t_1)_\Delta + (t_1 - t_2)_\Delta.        
    \end{split}
    \end{equation}
    The only case when the equation has a valid solution is when $t_1 = t_2$ and $t_3 \neq t_4$,
    since in this case the left hands side evaluates to $\Delta + 2$, while the right side evaluates to $2 \Delta$.\\
    Repeating the analysis for the fastest paths among $v_2$ and $v_4$,
    we get that the only valid solution is when $t_2 = t_3$ and $t_1 \neq t_4$.
    Altogether, we get that $f$ is not in a fastest path of length $2$ in $S_{u,v}$ if and only if the label of edge $f$ is the same as the labels on the edges incident to it, while the last remaining edge has a different label.
\end{proof}
